### 2018-12-16 PaperTitle.tex
Updates from ShareLaTeX

%%%% Proceedings format for most of ACM conferences (with the exceptions listed below) and all ICPS volumes.
\documentclass[sigconf]{acmart}
%%%% As of March 2017, [siggraph] is no longer used. Please use sigconf (above) for SIGGRAPH conferences.

%%%% Proceedings format for SIGPLAN conferences 
% \documentclass[sigplan, anonymous, review]{acmart}

%%%% Proceedings format for SIGCHI conferences
% \documentclass[sigchi, review]{acmart}

%%%% To use the SIGCHI extended abstract template, please visit
% https://www.overleaf.com/read/zzzfqvkmrfzn

\usepackage{booktabs} % For formal tables
\usepackage{graphicx}
\usepackage{subcaption}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

% Copyright
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

\graphicspath{{images/}}

\begin{document}

\title{The Next Release Problem in Real World}
\subtitle{Using Bugzilla dataset.}

\author{Young-kyoung Bae}
\affiliation{%
  \institution{KAIST, 20130298}
}
\email{tardis_95@kaist.ac.kr}

\author{Jin-woo Lee}
\affiliation{%
    \institution{KAIST, 20120774}
}
\email{hiall21@kaist.ac.kr}

\author{Sue-hwan Lee}
\affiliation{%
    \institution{KAIST, 20110686}
}
\email{wwa098@kaist.ac.kr}

\begin{abstract}
This report provides the work of applying NRP problems to real-world problem. The NRP problem is where the company chooses which requirements to reflect when releasing the next version. We criticize the fact that most of the previous papers extracted results using random variables rather than actual data values. To improve that, we've taken the data out of the Bugzilla in Firefox and Eclipse repositories, and we've solved the problem of choosing which bugs need to be fixed in the next version. We compared performance by applying several known algorithms, SA, GA NSGA2, IHP, and AHP. This report shows how well the application of search-based algorithms to NRP problems in the real world, and which algorithms are the most effective.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\keywords{NRP, MONRP, Bugzilla, Firefox, Eclipse}

\maketitle

\input{PaperBody.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{ref}

\end{document}