### 2018-12-17 PaperBody.tex
Updates from ShareLaTeX

\section{Introduction}
 The Next Release Problem(NRP) is the choice of what requirements the company will reflect when making the next version of the product. This problem is primarily aimed at helping the company get the most profit from limited costs. However, most of the previous studies were using random variables, not actual data values.

 The reason is that first of all, it's quite difficult to define the cost and the profit. Cost and profits are quite difficult to quantify. Both have to be calculated before the actual requirements are applied, so we have to use estimates value, rather than actual values. The second reason is that the requirements for the company's next production are highly secured data to the company and are very difficult to use in real research. Even if your company provides data, the actual results are prohibited to open, and this often results in post-processing after the release.

 We point out that these processes are wrong. The real NRP is that the cost and profit are not independent, and because it is used in the real industry, the process of obtaining the cost and profit has to be included to entire process. Also, importing and writing data from the future makes it difficult for research results to be applied to actual industrial sites. So we chose our topic to build our own industrial data and apply search-based algorithms.

Section 2 introduces the related paper about our topic. Section 3 introduces the basic model of the NRP problem that we are currently dealing with in our report.

We divided the main part of the report into two parts. First, in section 4, we will organize the requirements in the actual product and create a dataset. And in this process, we're not going to use any random variables, and we're not going to bring in actual profit and cost from the future. 
And in Section 5, we're going to apply different search-based algorithms to the datasets that we've created. In addition to SA, GA, and NSGA2, we also implemented ILH and AHP directly and applied them to datasets. We will apply them to actual datasets and compare the performance of each algorithm in terms of time cost and optimization answers. And we'll also see how much search-based algorithm has increased the profit compared to the actual true value.

All codes used in the paper can be found at

https://github.com/shkevinlee/CS454

\section{Related Work}
The Next Release Problem(NRP)\cite{NRP} was proposed by Bagnall et al in 2001. In this paper, authors considered requirements which can be dependent on other requirements. And also, they set the importance(or weight) of each customer who demands for resolving requirements to company.

In the Multi-Objective Next Release Problem(MONRP)\cite{MONRP}, authors tried to solve this problem as the multi-objective optimization problem. They applied algorithms(Pareto GA, Single-objective GA, and NSGA-II) and concluded that the NSGA-II is well suited to the MONRP.

After these papers, there are many trials to select the optimal set of requirements.\cite{ILP} \cite{DE} \cite{ACO} However, in almost cases, they used the data set which is randomly generated.

Tonella et al. were applied the Interactive Genetic Algorithm and AHP to the NRP and requirements prioritization for real software system as part of the project ACube, but their constraint was number of information about the priority graph, not the limited budget.\cite{IGA}

\section{The Model}

Let $R = \{r_1, r_2, \ldots, r_N\}$ is a set of all possible N requirements for software or company and each requirement $r_i \in R$ has its cost and profit, $cost_i \in \mathbb{Z}^+$ and $profit_i \in \mathbb{Z}^+$, to be performed. The cost is whatever company have to pay to resolve requirements, for example, money, time and etc. The profit is whatever company get by resolving requirements, for example, user satisfaction. In this project, we assumed that all requirements are independent. It means, there is no prerequisite of requirements. The cost vector $Cost$ and the profit vector $Profit$ for requirements are denoted by 
\[
Cost = \{cost_1, cost_2, \ldots, cost_N\}
\]
\[
Profit = \{profit_1, profit_2, \ldots, profit_N\}
\]

The company always wants to select requirements for the next release to get the highest sum of profits in the limited budget. Let the decision vector $\textbf{x} = \{x_1, x_2, \ldots, x_n\} \in \{0, 1\}$ where $x_i$ is 1 if requirement $i$ is selected and 0 otherwise. Then our objective function can be written as below:
\[
maximize \sum_{i = 1}^{N} profit_i \cdot x_i
\]
subject to
\[
\sum_{i = 1}^{N} cost_i \cdot x_i \leq B
\]
for some bound $B \in \mathbb{Z}^+$ which represents the limited budget.

\section{Data sets}

\begin{table} [H]
  \caption{Dataset from Bugzilla}
  \label{tab:commands}
  \begin{tabular}{cccccl}
    \toprule
    &ECLIPSE&FIREFOX-I&FIREFOX-II&FIREFOX-III\\
    \midrule
    Cost& 50699.70& 262945.50& 262945.50& 262945.50 \\
    Profit& 3489.64& 13025.88& 12210& 26332.88 \\
    \bottomrule
  \end{tabular}
\label{table:Dataset}
\end{table}

\subsection{Bugzilla}
\subsection{Open Source repository}
\subsection{Crawling Modeling}
\subsection{Research Question}
\subsection{Profit}
\subsection{Cost}

\section{Algorithms}
We applied two deterministic algorithms(AHP and ILP) and three heuristic algorithms(SA, GA and NSGA-II) which were introduced as the algorithms well-suited for the NRP in papers\cite{NRP}\cite{ILP}\cite{IGA}\cite{MONRP}. And, to encourage the performance of heuristic algorithms, we put a greedy solution into the initial population or initial decision vector.

The greedy algorithm for the initial solution of heuristic algorithms is described as below.

\begin{algorithm}
\caption{greedy algorithm}\label{alg:greedy}
\begin{algorithmic}
    \State Let $\textbf{x} = \{0, 0, \ldots, 0\}$ ;
    \State Let $CostSum = 0$ ;
    \For{$i = 1$ to $N$}
        \State $p = argmax(Profit)$ ;
        \If {$CostSum + cost_p \leq B$}
            \State $x_p = 1$ ;
            \State $CostSum = CostSum + cost_p$ ;
        \EndIf
        \State $profit_p = -\infty$ ;
    \EndFor
    \textbf{return} $x$
\end{algorithmic}
\end{algorithm}

\subsection{Analytic Hierarchy Process}
The Analytic Hierarchy Process(AHP) which was first applied to requirement prioritization by Karlsson and Ryan\cite{AHP} is a structured technique for organizing and analyzing complex decisions. To use the AHP for decision making in general, we had to perform pairwise comparisons of all the requirements to get their relative intensity. But, in our case, we don't have to compare relatively each other since each requirements has its own profit and cost. Therefore, we defined the relative intensity as the ratio of each requirement and made the profit-comparison matrix and the cost-comparison matrix by using them. And then, we calculated the priority vector. To get the optimal decision vector, we sorted the priority vector by its weight and selected requirements while the sum of costs didn't exceed the limited budget.

\subsection{Integer Linear Programming}
The Integer Linear Programming(ILP) is optimization method whose objective function and constraints are linear, and all variables are integer. The NRP can be treated as the ILP since each element of the decision vector $\textbf{x}$ is 0 or 1 and our constraint, the limited budget, is also linear. So we applied the ILP solver to the NRP problem.

Reference paper\cite{ILP} used ILP solver named \textit{CPLEX}, but it is commercial solver so we used \textit{PuLP}(modeler) and \textit{Gurobi}(solver) for this project. \textit{PuLP} is free open source software so users can use it. However, general \textit{Gurobi} solver is commercial one so we registered to \textit{Gurobi (www.gurobi.com)} and used \textit{Gurobi} academic version.

\subsection{Simulated Annealing}
Simulated annealing(SA) is one of the local search techniques which are iterative in nature and rely on the definition of a solution neighbourhood. For our NRP project, a requirement-based neighbourhood was defined. So the current solution can reach its neighbour by making small changes of decision vector $x$.

To apply the SA, we have to define a objective function to get the optimal decision vector:
\[
f(\textbf{x}) = \sum_{i = 1}^{N} profit_i \cdot x_i + \lambda \min \Big\{0, \sum_{i = 1}^{N} (cost_i \cdot x_i) - B \Big\}
\]
where $\textbf{x}$ is the decision vector, $f$ is our objective function and $\lambda$ is control parameter. We can impose a penalty if the sum of costs exceed the limited budget by using the control parameter $\lambda$. This objective function is also used in genetic algorithm to evaluate each chromosome.

In the NRP paper\cite{NRP}, the cooling schedule by Lundy and Mess\cite{LundySA} is well-suited. Their approach is presented below:
\[
    T_{i+1} = \frac{T_i}{1 + \beta T_i}
\]
where $T_i$ is the temperature at iteration $i$ and $\beta$ is control parameter. So for the Lundy and Mess cooling schedule, the temperature drops after each move.

The main loop of SA can be described in Algorithm \ref{alg:SA}. To increase the performance, we used a greedy solution as the initial decision vector.

In our project, the parameter $\lambda$ and $\beta$ were set to 5 and $1 \times 10^{-8}$ since Bagnall et al. got the best solution for this value in their paper.

\begin{algorithm}
\caption{Simulated Annealing (SA)}\label{alg:SA}
\begin{algorithmic}
    \State $x \gets$ initial decision vector ;
    \State $T \gets$ initial temperature ; 
    \While{not stopping condition}
        \State $x_{new} \gets GetRandomNeigbour(x)$ ;
        \If {$AcceptanceRate(F(x), F(x_{new}), T) \geq random(0, 1)$}
            \State $x \gets x_{new}$ ;
        \EndIf
        \State $T \gets COOL(T)$ \Comment{Lundy and Mess}
    \EndWhile
    \textbf{return} $x$
\end{algorithmic}
\end{algorithm}

\subsection{Genetic Algorithm}
The Genetic Algorithm(GA) is one of the evolutionary algorithms. In our project, we used a single objective function for GA which is the same with the simulated annealing. To make offspring population, we used the tournament selection, cross-over, and mutation operator. We also let the best one which has the best profit remain on the population for elitism.

\subsection{NSGA-II}
The Non-dominated Sorting Genetic Algorithm-II (NSGA-II) was introduced by Deb et al.\cite{NSGA2} It is one of the various approaches to solve the Multi-Objective Optimization Problems(MOOPs). So, to apply this algorithm, we have to decide the second objective function.

The goal of our project is finding the optimal decision vector under the limited budget. Therefore, if we convert this single objective function into the MOOPs, our objective functions on MOOPs can be written as below:

\[
maximize \sum_{i = 1}^{N} profit_i \cdot x_i
\]
\[
maximize -\sum_{i = 1}^{N} cost_i \cdot x_i
\]

The result of NSGA-II is the Pareto front on two objective function space. So, to get the optimal decision vector among them, we selected one which has maximum profits under the limited budget among the Pareto front.

The main loop of NSGA-II can be described in Algorithm \ref{alg:nsga2}. Initial population is randomly created except one which is greedy solution. The \textit{make-new-pop} function consist of tournament selection, cross-over, and mutation operator which are used to create a child population.

\begin{algorithm}
\caption{NSGA-II}\label{alg:nsga2}
\begin{algorithmic}
    \State $t = 0$ ;
    \State $P_0 \gets$ initial population ;
    \State $Q_0 =$ make-new-pop($P_0$) ;
    \While{not stopping condition}
        \State Let $R_t = P_t \cup Q_t$ ;
        \State Let $F = $ fast-non-dominated-sort ($R_t$) ; %\Comment{}
        \State Let $P_{t+1} = \emptyset$ and $i = 1$ ;
        \While{$|P_{t+1}| + |F_i| \leq N$}
            \State Apply crowding-distance-assignment($F_i$) ; %\Comment{}
            \State Let $P_{t+1} = P_{t+1} \cup F_i$ ;
            \State Let $i = i + 1$
        \EndWhile
        \State $Sort(F_i, \prec_n)$
        \State Let $P_{t+1} = P_{t+1} \cup F_i[1:(N - |P_{t+1}|)]$ ;
        \State Let $Q_{t+1} = $ make-new-pop($P_{t+1}$) ;
        \State Let $t = t + 1$ ;
    \EndWhile
\end{algorithmic}
\end{algorithm}

%\begin{table*}
%  \caption{Profit Comparison}
%  \label{tab:commands}
%  \begin{tabular}{cccccl}
%    \toprule
%    &eclipse&firefox-I&firefox-II&firefox-III\\
%    \midrule
%    Estimated Cost&50699.70190&262945.49581&262945.49581&262945.49581 \\
%    Real Profit&3489.63592&13025.88085&12210&26332.88085 \\
%    \hline
%    SA&3702.85182&13524.93411&16300.0&28358.2295685 \\
%    GA&3793.5045&13662.38600&16303.0&28522.11861 \\
%    NSGA-II&3822.47741&13700.52067&16332.6667&28729.33476 \\
%    ILP& \textbf{3825.42768}&\textbf{13768.44722}&\textbf{16336.0}&\textbf{28%845.23124} \\
%    AHP&3810.3125825&13758.5577962&16249.0&28811.2749907 \\
%    \bottomrule
%  \end{tabular}
%\label{table:DatasetProfit}
%\end{table*}
% end the environment with {table*}, NOTE not {table}!

\begin{figure*}[h]
\begin{subfigure}{\columnwidth}
\includegraphics[width=0.9\linewidth]{bugzilla_eclipse_log(comments)_2016meancost.JPG}
\caption{ECLIPSE dataset. \normalfont{The estimated cost is 50699.70190}}
\end{subfigure}
\begin{subfigure}{\columnwidth}
\includegraphics[width=0.9\linewidth]{bugzilla_firefox_log(comments)_2016meancost.JPG}
\caption{FIREFOX-I. \normalfont{The estimated cost is 262945.49581}}
\end{subfigure}
\begin{subfigure}{\columnwidth}
\includegraphics[width=0.9\linewidth]{bugzilla_firefox_priority_2016meancost.JPG}
\caption{FIREFOX-II. \normalfont{The estimated cost is 262945.49581}}
\end{subfigure}
\begin{subfigure}{\columnwidth}
\includegraphics[width=0.9\linewidth]{bugzilla_firefox_comments-priority_2016meancost.JPG}
\caption{FIREFOX-III. \normalfont{The estimated cost is 262945.49581}}
\end{subfigure}

\caption{Profit Comparison for each dataset}
\label{fig:DatasetProfit}
\end{figure*}

%\begin{figure*}[h]
%\begin{subfigure}{\columnwidth}
%\includegraphics[width=0.9\linewidth]{SA-correlation.JPG}
%\caption{SA}
%\label{fig:subim5}
%\end{subfigure}
%\begin{subfigure}{\columnwidth}
%\includegraphics[width=0.9\linewidth]{GA-correlation.JPG}
%\caption{GA}
%\label{fig:subim6}
%\end{subfigure}
%\begin{subfigure}{\columnwidth}
%\includegraphics[width=0.9\linewidth]{ILP-correlation.JPG}
%\caption{ILP}
%\label{fig:subim7}
%\end{subfigure}
%\begin{subfigure}{\columnwidth}
%\includegraphics[width=0.9\linewidth]{AHP-correlation.JPG}
%\caption{AHP}
%\label{fig:subim8}
%\end{subfigure}

%\caption{Cost-Profit Correlation}
%\label{fig:image2}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=0.9\linewidth]{ParetoFronts.png}
%\caption{Cost-Profit Correlation in one graph}
%\label{fig:paretofronts}
%\end{figure*}

\section{Experimental Set up}
In this section, we describe two experiments for comparing algorithms on datasets and analyzing correlation between cost and profit. We used four datasets(\textit{ECLIPSE, FIREFOX-I, FIREFOX-II} and \textit{FIREFOX-III}) from Bugzilla on experiments which were described in \textbf{section 4}.

At first experiment(EX1), we applied each algorithm to four datasets in order to compare their performances. And we hope to know the effect by different profit strategies through analyzing \textit{FIREFOX-I}, \textit{FIREFOX-II}, and \textit{FIREFOX-III} dataset.

At second experiment(EX2), we applied each algorithm to \textit{FIREFOX-III} dataset to analyze the correlation between cost and profit. And also, we expected to get the interesting range which have a great change as the limited budget are increased or decreased.

All heuristic approaches(SA, GA and NSGA-II) were run for 500,000 function evaluations. SA started from a greedy solution and GA and NSGA-II started from the initial population which contains a greedy solution. The population size was set to 200. So each experimental execution of GA and NSGA-II was terminated when the generation number reached 2,500. GA and NSGA-II used the tournament selection(tournament size is 5), single-point cross-over, and bit-wise mutation for binary-coded GAs. The cross-over probability and the mutation probability were set to 0.8 and 1/N (where N is the length of chromosome same with the number of requirements). Every heuristic algorithms were executed 3 times for each experiment.

\section{Result and Analysis}

%At first experiment, we applied five algorithms to four datasets and compared the maximum profit each other. In \textbf{Table \ref{table:DatasetProfit}}, upper two rows are estimated cost sum and real profit sum of each datasets. Below these rows, there are averaged profits for each algorithms on four datasets. At the result, every profits for algorithms are bigger than real profit sum. Also, consumed time to return maximum profit was exceedingly short in ILP and AHP, rather than SA, GA, and NSGA-II.

%\textbf{Figure \ref{fig:DatasetProfit}} is the graphs of \textbf{table 1}. Each graph %expresses the averaged profit of five algorithms. At the result, the profit of ILP is %the biggest, and the profit of NSGA-II and AHP is bigger than SA and GA in most case. %Except ILP and AHP, maximum profit is high in order of NSGA-II, GA, then SA.

%At experiment 2, we got the profits while changing the cost constraint. The result is expressed in \textbf{Figure \ref{fig:paretofronts}}. \textbf{Figure \ref{fig:paretofronts}} expresses maximum profits on each cost, and \textbf{Figure 3} is integrating graph of \textbf{Figure 2} because of comparing the difference for four algorithms.

%The result is that they are all increasing almost similarly. One of the reasons to perform this experiment was that we wondered if there is a peculiar part in cost-profit graph.(if there is exceeding increase in profit) And at the result, there is no prominently peculiar part, and all the results are increasing normally.

\subsection{Experiment 1}
At EX1, we applied two deterministic and three heuristic algorithms to four datasets and compared the averaged maximum profit to each other. Figure \ref{fig:DatasetProfit} shows the results of EX1. Each graph expresses the profit of five algorithms under the limited budget(= the estimated cost).

In these results, the profit of each algorithm were bigger than the real profit in Table \ref{table:Dataset}. And, in every case, the ILP has the biggest profit. However, the order of other algorithms are slightly changed for dataset.

The SA has the smallest profit in most cases, except Figure \ref{fig:DatasetProfit}-(c). Although the SA started from a greedy solution, there is little difference between the result and a greedy solution. The profit of GA is bigger than SA and smaller than NSGA-II in every cases. So, the profit is high in order of NSGA-II, GA, then SA. The AHP has good performances for some cases. However, it has the lowest profit in \ref{fig:DatasetProfit}-(c). Therefore, we can conclude that the strategy about evaluating profit can be affect the results.

The run time of two deterministic algorithms were much faster than heuristic algorithms. Therefore, we can conclude that the ILP is well-suited for our datasets and also it is much faster than other heuristic algorithms.

\subsection{Experiment 2}
\begin{table*}[]
\caption{Profit on each limited budget}
\begin{tabular}{@{}crrrrr@{}}
\toprule
\multicolumn{1}{c}{\multirow{2}{*}{\begin{tabular}[c]{@{}c@{}}The limited\\ budget\end{tabular}}} & \multicolumn{5}{c}{Algorithms}                                                                                                    \\ \cmidrule(l){2-6} 
\multicolumn{1}{c}{}                                                                              & \multicolumn{1}{c}{AHP} & \multicolumn{1}{c}{ILP} & \multicolumn{1}{c}{SA} & \multicolumn{1}{c}{GA} & \multicolumn{1}{c}{NSGA-II} \\ \midrule
100,000 & 14472.21303& \textbf{14530.92923}& 13787.94014& 13916.62496& 14359.51809\\
120,000 & 16798.95935& \textbf{16845.48739}& 16269.28033& 16440.60658& 16746.14339\\
140,000 & 18948.34660& \textbf{18979.35023}& 18553.36746& 18706.04924& 18906.95215\\
160,000 & 20906.31785& \textbf{20934.94744}& 20559.69749& 20656.69427& 20873.08535\\
180,000 & 22683.10701& \textbf{22718.92665}& 22332.52554& 22449.04724& 22651.51132\\
200,000 & 24343.18852& \textbf{24369.86106}& 23894.25792& 24054.60654& 24281.04096\\
220,000 & 25881.18360& \textbf{25901.23028}& 25425.18545& 25565.45349& 25804.51079\\
240,000 & 27305.09756& \textbf{27324.51067}& 26790.58789& 26987.08126& 27214.27095\\
260,000 & 28626.58769& \textbf{28653.54334}& 28161.56720& 28293.21833& 28556.63614\\ \bottomrule
\end{tabular}
\label{table:ParetoFront}
\end{table*}

At EX2, we got the profits while changing the cost constraint. The result is expressed in Table \ref{table:ParetoFront}. In the results, profits of each algorithms are all increasing almost similarly and the order of profits(ILP, AHP. NSGA-II, GA, then SA) are same in every case. And also, we observed that the change of profits become smaller as the limited budget increases.

One of the reasons to perform EX2 was that we wondered if there is a peculiar part in cost-profit graph. However, we cannot observe prominently peculiar part in the dataset.

\section{Conclusion}
The research question we presented was two things.
The RQ1. "Could we estimate profit from Bugzilla without forward looking?". We were able to get a proposal for the demand of custmers in Bugzilla through Priority and Comment. The RQ2. "Could we estimate profit from Bugzilla without forward looking?". We define the Cost in Bugzilla as time. The component's Cost is not accurate. However, the costs are not meaningless given the trends of 2016 and 2017. With this research question, we have built a dataset in the real world for NRP research.

In the results of the experiment.
EX1. was profit comparison. Exept for dataset with priority, the ILP has the highest result. And deterministic algorithms that AHP and ILP are very fast than heuristic algorithms. Ex2. is a comparison of the results of the profit according to the cost limit. As we mentioned, we found that limits of the Cost and the Profit had a linear relationship, regardless of algorithm. These results show that tere are no major problems with applying NRP argorithms to our dataset. Therefore, the dataset was well set to solve  the NRP.

In previous NRP studies, both dataset were applied with the random Cost and Profit values. But our project use actual situations value. It also showed that the NRP's algorithm worked well. In this way, later NRP studies have shown the possibility of using real situations rather than random values.